---
title: "STAT 350 Final Project - Yan & Leo"
author: "Abs_orbo"
date: "10/24/2020"
output: pdf_document
---
Question: Regression analysis on US 2000 death rate

Project Information
```{r}
# we are assigned Dataset 1 (Cancer)
# https://data.world/nrippner/cancer-linear-regression-model-tutorial


# The	following	sections	must	be	included:
#    Abstract (150	words	or	less)
#    Introduction (must	contain	a	thorough	description	of	the	questions	of	interest)
#    Data	Description (must	contain	data	visualizations	that	are	properly	labelled	and	explained)
#    Methods (must	contain	a	complete	description	of	all	analysis	tools	used)
#    Results (all	figures	should	be	properly	labelled	and	discussed)
#    Conclusion (must	contain	a	concise	discussion	of	what	has	been	learned	about	the	application	from	the	analysis)
#    Appendix (must	include	all	data	and	R	Markdown	files	for	reproducibility)


# BEFORE commencing	your	analysis, you	must	introduce	one	new	additional	data	point into	your	assigned	dataset.	A	description	of	this	unique data	point	must	be	included	in	your	Data	 Description section	along	with	some rationale	for	the	values	chosen.		


# Grading	Scheme:
# 8 Overall	presentation	and	organization	of	materials (includes	quality	of	writing	and	quality	of	
# the	Github	presentation,	…)
# 5 Quality	of	data	visualizations
# 8 Correctness	of	analysis
# 6 Quality	and	selection	of	relevant	figures	
# 8 Interpretation	of	results
```

```{r "setup", include=FALSE}
knitr::opts_knit$set(root.dir = "C:\\Users\\LYC\\Desktop\\SFU\\Fall 2020 Courses\\STAT 350\\Final project\\data")
```

Initialization
```{r}

```

Import datasets
```{r}
death_rate = read.csv("death.csv")
b1_population = read.csv("cc00_tab_B1_area_population.csv")
b2_pop_by_age_sex_race = read.csv("cc00_tab_B2_pop_by_age_sex_race.csv")
b4_vital_stat_health = read.csv("cc00_tab_B4_vital_statistics_and_health.csv")
b5_education_income_poverty = read.csv("cc00_tab_B5_education_income_poverty.csv")
b6_crime_housing = read.csv("cc00_tab_B6_crime_housing_building_permit.csv")
b7_labor_employment = read.csv("cc00_tab_B7_labor_force_employment.csv")
b8_income = read.csv("cc00_tab_B8_personal_income_earnings.csv")
b13_govt_programs = read.csv("cc00_tab_B13_govt_programs_employment_finances.csv")
```

Data preprocessing
```{r}
death_rate = b4_vital_stat_health$B4.vst05
FIPS = b1_population$B1.geo01
county_name = b1_population$B1.geo09
population = b1_population$B1.pop03

num_care_facilities = b4_vital_stat_health$B4.hlt06
per_below_poverty = b5_education_income_poverty$B5.pov04
per_high_grad = as.numeric(b5_education_income_poverty$B5.edc06) - as.numeric(b5_education_income_poverty$B5.edc08)
per_bachelor_grad = b5_education_income_poverty$B5.edc08
num_violent_crimes = b6_crime_housing$B6.crm02
num_property_crimes = b6_crime_housing$B6.crm03
unemploy_rate = b7_labor_employment$B7.clf05
per_cap_income_as_per_national_average = b8_income$B8.pie06
num_social_security_ben_per_1000_resid = b13_govt_programs$B13.soc04

X = cbind(death_rate,FIPS,county_name,population,num_care_facilities,per_below_poverty,per_high_grad,per_bachelor_grad,num_violent_crimes,num_property_crimes,unemploy_rate,per_cap_income_as_per_national_average,num_social_security_ben_per_1000_resid)

X = as.data.frame(X)

states = toupper(c("Alabama","Alaska","Arizona","Arkansas","California","Colorado","Connecticut","Delaware","Florida","Georgia","Hawaii","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts","Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada","New Hampshire","New Jersey","New Mexico","New York","North Carolina","North Dakota","Ohio","Oklahoma","Oregon","Pennsylvania","Rhode Island","South Carolina","South Dakota","Tennessee","Texas","Utah","Vermont","Virginia","Washington","West Virginia","Wisconsin","Wyoming"))

# delete US row
X = X[-1,]

#delete State rows
for (i in states){
  X = X[!grepl(i,X$county_name),]
}

# convert columns to numeric
i = c(1,4,5,6,7,8,9,10,11,12,13)

X[,i] = apply(X[,i], MARGIN = 2, function(x) as.numeric(as.character(x)))

# delete rows with NA values
X = X[-which(is.na(X), arr.ind = TRUE),]

# feature scaling (standardize)
X_scaled = X
X_scaled[,i] = scale(X_scaled[,i])

```

Abstract (150 words or less)
```{r}

```

Introduction (must contain a thorough description of the questions of interest)
```{r}

```

Data Description (must contain data visualizations that are properly labelled and explained)
```{r}

```

Methods (must contain a complete description of all analysis tools used)
```{r}

```

Results (all figures should be properly labelled and discussed)
```{r}

```

Conclusion (must contain a concise discussion of what has been learned about the application from the analysis)
```{r}

```

Appendix (must include all data and R Markdown files for reproducibility)
```{r}

```

