---
title: "STAT 350 Final Project - Yan & Leo"
author: "Abs_orbo"
date: "10/24/2020"
output: pdf_document
---
Project Information
```{r}
# we are assigned Dataset 1 (Cancer)
# https://data.world/nrippner/cancer-linear-regression-model-tutorial


# The	following	sections	must	be	included:
#    Abstract (150	words	or	less)
#    Introduction (must	contain	a	thorough	description	of	the	questions	of	interest)
#    Data	Description (must	contain	data	visualizations	that	are	properly	labelled	and	explained)
#    Methods (must	contain	a	complete	description	of	all	analysis	tools	used)
#    Results (all	figures	should	be	properly	labelled	and	discussed)
#    Conclusion (must	contain	a	concise	discussion	of	what	has	been	learned	about	the	application	from	the	analysis)
#    Appendix (must	include	all	data	and	R	Markdown	files	for	reproducibility)


# BEFORE commencing	your	analysis, you	must	introduce	one	new	additional	data	point into	your	assigned	dataset.	A	description	of	this	unique data	point	must	be	included	in	your	Data	 Description section	along	with	some rationale	for	the	values	chosen.		


# Grading	Scheme:
# 8 Overall	presentation	and	organization	of	materials (includes	quality	of	writing	and	quality	of	
# the	Github	presentation,	…)
# 5 Quality	of	data	visualizations
# 8 Correctness	of	analysis
# 6 Quality	and	selection	of	relevant	figures	
# 8 Interpretation	of	results
```

Abstract (150 words or less)
```{r}

```

Introduction (must contain a thorough description of the questions of interest)
```{r}

```

Data Description (must contain data visualizations that are properly labelled and explained)
```{r}

```

Methods (must contain a complete description of all analysis tools used)
```{r}

```

Results (all figures should be properly labelled and discussed)
```{r}

```

Conclusion (must contain a concise discussion of what has been learned about the application from the analysis)
```{r}

```

Appendix (must include all data and R Markdown files for reproducibility)
```{r}

```

